---
title: "Take-home Exercise 2"
description: Creating data visualisation beyond default
author: Khoo Wei Lun
date: "January 29, 2023"
toc: true
toc-location: left
---

# 1. Overview

In this DataViz Makeover exercise, we are going to critique a

# 2. Original Design

2.1 Clarity

2.2 Aesthetics

# 3. Alternative Design

3.1 Proposed design sketch

3.2 Design principles and best practices implemented

# 4. Loading Libraries and Dataset

The following code chunk is used to install the required R packages and load them onto R environment.

```{r}
pacman::p_load(patchwork, ggthemes, hrbrthemes, ggrepel, tidyverse)
```

Next, read_csv from readr package is used to import the source file which is in csv file format.

```{r}
popdata <- read_csv("Data/respopagesextod2022.csv")
glimpse(popdata)
```

# 5. Data Preparation

**5.1 Reformatting incorrect data type**

The function factor() from Base R is used to convert variables Age Group (AG) and Planning Area (PA) to factors.

```{r}
#pop$AG <- factor (pop$AG, levels = unique(pop$AG))
#pop$PA <- factor (pop$PA, levels = unique(pop$PA))
```

**5.2**

Male population values need to be

plot the male generation in opposite direction

References: [R4DSA meetup - R Series Episode 2: Doing Data Science with Rtidyverse methods (r4dsa-meetup.netlify.app)](https://r4dsa-meetup.netlify.app/smulibr/episode02-tidyverse_methods#/data-science-process---an-overview)

# 6. Visualisation

**6.1 Finding the top 9 most populated planning area**

6.2 Trellis diagram
