---
title: "In-Class Excerise 5"
date: "February 11, 2023"
date-modified: "`r Sys.Date()`"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

# Installing and Launching R Packages

```{r}
pacman::p_load(seriation, dendextend, heatmaply, GGally, parallelPlot, tidyverse)
```

# Importing and Preparing The Data

------------------------------------------------------------------------

Importing data set

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

Preparing the data

```{r}
row.names(wh) <- wh$Country
```

Transforming the data frame into matrix

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

# Heatmap

------------------------------------------------------------------------

Plotting static heatmap

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

Creating interactive heatmap

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

# Clustering

------------------------------------------------------------------------

Plotting heatmap by using hierachical clustering algorithm with "Euclidean distance" and "ward.D" method.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

*dend_expend()* will be used to determine the recommended clustering method to be used.

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

*find_k()* is used to determine the optimal number of cluster

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

Visualising the clustering

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

Using Blue color paletter of rColorBrewer

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

Adding plotting features to ensure cartographic quality:

-   *k_row* is used to produce 5 groups.

-   *margins* is used to change the top margin to 60 and row margin to 200.

-   *fontsizw_row* and *fontsize_col* are used to change the font size for row and column labels to 4.

-   *main* is used to write the main title of the plot.

-   *xlab* and *ylab* are used to write the x-axis and y-axis labels respectively.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```

# Parallel Coordinates Plot

------------------------------------------------------------------------

**Plotting basic static parallel coordinates**

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

**Adding boxplot**

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

Learning points:

-   `groupColumn` argument is used to group the observations (i.e. parallel lines) by using a single variable (i.e. Region) and colour the parallel coordinates lines by region name.

-   `scale` argument is used to scale the variables in the parallel coordinate plot by using `uniminmax` method. The method univariately scale each variable so the minimum of the variable is zero and the maximum is one.

-   `alphaLines` argument is used to reduce the intensity of the line colour to 0.2. The permissible value range is between 0 to 1.

-   `boxplot` argument is used to turn on the boxplot by using logical `TRUE`. The default is `FALSE`.

**Parallel coordinates with facet**

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

**Plotting Interactive Parallel Coordinates Plot**

Basic plot

```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

Adding histogram

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```
